Variational Inference for Inverse Reinforcement Learning with Gaussian Processes

### 2018-11-04 mprop/mprop.tex
Set Theta = lambda, and explain the MDP representation a little better

\newcommand{\pfull}{p(\mathcal{D}, \Theta, \mathbf{X_u}, \mathbf{u}, \mathbf{r})}
\newcommand{\pfullS}{p(\mathcal{D}, \Theta_s, \mathbf{X_u}, \mathbf{u}_s, \mathbf{r}_s)}
\newcommand{\posterior}{p(\Theta, \mathbf{u}, \mathbf{r} | \mathcal{D}, \mathbf{X_u})}
\newcommand{\approximation}{q_{\bm\nu}(\Theta, \mathbf{u}, \mathbf{r})}
\newcommand{\approximationS}{q_{\bm\nu}(\Theta_s, \mathbf{u}_s, \mathbf{r}_s)}

\newcommand{\Eq}{\mathbb{E}_{(\Theta, \mathbf{u}, \mathbf{r}) \sim \approximation}}

@@

\section{Introduction}
%briefly explain the context of the project problem
% TODO: introduce a data set
% TODO: example application for motivation (why would the world be better)
% TODO: (possibly) a figure with my notation
% TODO: explain the MDP Q/V stuff
% TODO: introduce research questions

@@

In \emph{inverse reinforcement learning}, one is presented with an MDP without a
reward function $\mathcal{M} \setminus \{ r \}$ and a set of expert
demonstrations $\mathcal{D} = \{ \zeta_i \}_{i=1}^N$, where each demonstration
$\zeta_i = \{ (s_{i,0}, a_{i,0}), \dots, (s_{i,T}, a_{i,T}) \}$ is a multiset of
state-action pairs representing the actions taken by the expert during a
particular recorded session. Each state is also characterised by a number of
features. The goal of IRL is then to find $r$ such that the optimal policy under
$r$
\[ \pi^* = \argmax_\pi \mathbb{E}\left[ \sum_{t=0}^\infty \gamma^t r(s_t) | \pi
  \right] \]
matches the actions in $\mathcal{D}$.

The likelihood of the data can be written down as
\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}
\begin{equation} \label{pDr}
  p(\mathcal{D} | r) = \prod_{i=1}^N \prod_{t=1}^T p(a_{i,t} | s_{i,t}) = \exp\left( \sum_{i=1}^N \sum_{t=1}^T Q_r(s_{i,t}, a_{i,t}) - V_r(s_{i,t}) \right),
\end{equation}
where
\begin{equation}
  Q_r(s_{i,t}, a_{i,t}) = r(s_{i,t}) + \gamma\sum_{s' \in \mathcal{S}}
  \mathcal{T}(s_{i,t}, a_{i,t}, s')V_r(s'),
\end{equation}
and $V_r(s)$ can be obtained by repeatedly applying the equation
\cite{supplementary_material}
\[ V_r(s) = \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \gamma\sum_{s' \in
      \mathcal{S}} \mathcal{T}(s, a, s')V_r(s') \right). \]
However, a reward function learned by maximising this likelihood is not
transferable to new situations
\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}. One needs to model the
reward structure in a way that would allow reward predictions for previously
unseen states.

One way to model rewards without assumptions of linearity is with a
\emph{Gaussian process} (GP). A GP is a collection of random variables, any
finite combination of which has a joint Gaussian distribution
\cite{DBLP:books/lib/RasmussenW06}. We write $r \sim \mathcal{GP}(0,
k_{\Theta})$ to say that $r$ is a GP with mean $0$ and covariance function
$k_{\Theta}$, which uses a set of hyperparameters $\Theta$. Covariance functions
take two state feature vectors as input and quantify how similar the two states
are, in a sense that we would expect them to have similar rewards.

% TODO: positive definite
As training a GP with $n$ data points has a time complexity of
$\mathcal{O}(n^3)$ \cite{DBLP:books/lib/RasmussenW06}, numerous approximation
methods have been suggested, many of which select a subset of data called
\emph{inducing points} and focus most of the training effort on them
\cite{DBLP:journals/corr/abs-1807-01065}. Let $\mathbf{X_u}$ be the matrix of
features at inducing states, $\mathbf{u}$ the rewards at those states, and
$\mathbf{r}$ a vector with $r(\mathcal{S})$ as elements. Then the full joint
probability distribution can be factorised as
\begin{equation} \label{full}
  \pfull = p(\mathbf{X_u}) \times p(\Theta | \mathbf{X_u}) \times p(\mathbf{u}
  | \Theta, \mathbf{X_u}) \times p(\mathbf{r} | \Theta, \mathbf{X_u},
  \mathbf{u}) \times p(\mathcal{D} | r).
\end{equation}
Here $p(\mathbf{X_u})$ and $p(\Theta | \mathbf{X_u})$ are customisable priors,
\begin{align}
  p(\mathbf{u} | \Theta, \mathbf{X_u}) &= \mathcal{N}(\mathbf{u}; \mathbf{0}, \Kuu) \nonumber \\
                                       &= \frac{1}{(2\pi)^{m/2}|\Kuu|^{1/2}}\exp \left( -\frac{1}{2} \mathbf{u}^\intercal\Kuu^{-1}\mathbf{u} \right) \nonumber \\
                                       &= \exp\left(-\frac{1}{2}\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u} - \frac{1}{2}\log|\Kuu| - \frac{m}{2}\log 2\pi\right) \label{eq:normal}
\end{align}
is the GP prior \cite{DBLP:books/lib/RasmussenW06}, where $m \in \mathbb{N}$ is
the number of inducing points. The GP posterior is a
multivariate Gaussian \cite{DBLP:conf/nips/LevinePK11}
\begin{equation} \label{eq:r}
  p(\mathbf{r} | \Theta, \mathbf{X_u}, \mathbf{u}) =
  \mathcal{N}(\mathbf{r}; \Kru^\intercal\Kuu^{-1}\mathbf{u}, \Krr - \Kru^\intercal\Kuu^{-1}\Kru),
\end{equation}
and $p(\mathcal{D} | r)$ is as in \eqref{pDr}. The matrices such as
$\Kru$ are called \emph{covariance matrices} and are defined as
$[\Kru]_{i,j} = k_\Theta(\mathbf{x}_{\mathbf{r},i}, \mathbf{x}_{\mathbf{u},j})$,
where $\mathbf{x}_{\mathbf{r},i}$ and $\mathbf{x}_{\mathbf{u},j}$ denote feature
vectors for the $i$th state in $\mathcal{S}$ and the $j$th state in
$\mathbf{X_u}$, respectively \cite{DBLP:conf/uai/JinDAS17}.

% TODO: where does the lower bound come from?
Given this model, data $\mathcal{D}$, and inducing feature matrix
$\mathbf{X_u}$, our goal is then to find optimal values of hyperparameters
$\Theta$, inducing rewards $\mathbf{u}$, and the reward function $r$. While the
previous paper that considered this IRL model computed maximum likelihood estimates
for $\Theta$ and $\mathbf{u}$, and made an assumption that $\mathbf{r}$ in
\eqref{eq:r} has zero variance \cite{DBLP:conf/nips/LevinePK11}, we aim to
avoid this assumption and use variational inference to approximate the full
posterior distribution $\posterior$. \emph{Variational inference} (VI) is an
approximation technique for probability densities \cite{blei2017variational}.
Let $\approximation$ be our approximating family of probability distributions
for $\posterior$ with its own hyperparameter vector $\bm\nu$. Then it is up
to VI algorithms to optimise $\bm\nu$ in order to minimise the
\emph{Kullback-Leibler} (KL) divergence between the original probability
distribution and our approximation.  KL divergence (asymmetrically) measures how
different the two distributions are and in this case can be defined as follows
\cite{blei2017variational}:
\[ \DKL(\approximation || \posterior) = \Eq \left[ \log
  \frac{\approximation}{\posterior} \right]. \]
Since KL divergence is typically hard to compute, instead of minimising it, VI
typically tries to maximise the \emph{evidence lower bound} (ELBO) defined as
\cite{DBLP:books/lib/Bishop07,blei2017variational}
\begin{equation} \label{eq:elbo}
  \begin{split}
    \mathcal{L}(\bm\nu) &= \Eq \left[ \log \frac{\pfull}{\approximation}
    \right] \\
    &= \iiint \approximation \log
    \frac{\pfull}{\approximation}\,d\Theta\,d\mathbf{u}\,d\mathbf{r}.
  \end{split}
\end{equation}

@@

\begin{table}
  \centering % TODO: recheck
  \begin{tabular}{llllll}
    \toprule
    Authors, year & Inducing points & Hyperparameters & Complexity \\
    \midrule
    Titsias, 2009 \cite{DBLP:journals/jmlr/Titsias09} & variational & variational & $\mathcal{O}(nm^2)$ \\
    Hensman et al., 2013 \cite{DBLP:conf/uai/HensmanFL13} & fixed & variational & $\mathcal{O}(m^3)$ \\
    Gal et al., 2014 \cite{DBLP:conf/nips/GalWR14} & variational & variational & $\mathcal{O}(nm^2)$ \\
    Hoang et al., 2015 \cite{DBLP:conf/icml/HoangHL15} & fixed & fixed & $\mathcal{O}(m^3)$ \\
    Cheng and Boots, 2017 \cite{DBLP:conf/nips/ChengB17} & variational & variational & $\mathcal{O}(nm_\alpha + nm_\beta^2)$ \\
    Hensman et al., 2017 \cite{DBLP:journals/jmlr/HensmanDS17} & fixed & variational & $\mathcal{O}(nm)$ \\
    Peng et al., 2017 \cite{DBLP:conf/icml/PengZZQ17} & variational & variational & $\mathcal{O}(m^3)$ \\
    \bottomrule
  \end{tabular}
  \caption{Summary of relevant VI approximations to GPs. For both inducing
    points and hyperparameters, 'fixed' means 'chosen before the algorithm
    starts' and 'variational' means 'included amongst the variational
    parameters'. Hyperparameters $m_\alpha$ and $m_\beta$ refer to the number of
    bases used to represent the GP's mean and covariance, respectively.}
  \label{table:approximations}
\end{table}

@@

%state how you propose to solve the software development problem. Show that your
%proposed approach is feasible, but identify any risks.
% TODO: Theta should be equal to lambda

In order to properly investigate the difference between variational inference
and maximum likelihood estimation for the model, we keep other parts of the
model the same. Namely, we set the covariance function to a version of the
automatic relevance detection kernel
\cite{DBLP:conf/nips/LevinePK11,neal2012bayesian}
\[ k_\Theta(\mathbf{x}_i, \mathbf{x}_j) = \lambda_0\exp\left(
    -\frac{1}{2}(\mathbf{x}_i - \mathbf{x}_j)^\intercal\bm\Lambda(\mathbf{x}_i -
    \mathbf{x}_j) - \mathbbm{1}[i \ne j]\sigma^2\Tr[\bm\Lambda] \right), \]
where $\lambda_0$ is the overall ``scale'' factor for how similar or distant the
states are,
\[ \bm\Lambda = \diag(\lambda_1, \dots, \lambda_d) \]
is a diagonal matrix that determines how relevant each feature is (where $d$
denotes the number of features), $\mathbbm{1}$ is defined as
\[ \mathbbm{1}[b] = \begin{cases}
    1 & \mbox{if $b$ is true} \\
    0 & \mbox{otherwise},
  \end{cases} \]
and $\sigma^2$ is set to $10^{-2}/2$ (as the original paper noted that
the value makes little difference to the performance of the algorithm
\cite{DBLP:conf/nips/LevinePK11}). Thus we can set $\bm\lambda = (\lambda_0,
\dots, \lambda_d)^\intercal$, and $\Theta = \{ \bm\lambda \}$. Similarly, we
keep the expression for the prior of $\Theta$:
\begin{equation} \label{eq:theta_prior}
  p(\Theta | \mathbf{X_u}) = \exp\left( -\frac{1}{2}\Tr[\Kuu^{-2}] -
    \sum_{i=1}^d \log(\lambda_i + 1) \right).
\end{equation}

Next, we can rewrite the posterior by using the chain rule and Bayes' theorem in
order to get a better sense of what we are trying to approximate:
\[
  \begin{split}
    \posterior &= p(\Theta | \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{u} |
    \Theta, \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{r} | \Theta,
    \mathbf{X_u}, \mathbf{u}, \mathcal{D}) \\
    &\propto p(\Theta | \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{u} | \Theta,
    \mathbf{X_u}, \mathcal{D}) \times p(\mathcal{D} | r) \times p(\mathbf{r} |
    \Theta, \mathbf{X_u}, \mathbf{u}) \\
    &\propto p(\Theta | \mathbf{X_u}, \mathcal{D}) \times p(\mathcal{D} |
    \Theta, \mathbf{X_u}, \mathbf{u}) \times p(\mathbf{u} | \Theta,
    \mathbf{X_u}) \times p(\mathcal{D} | r) \times p(\mathbf{r} | \Theta,
    \mathbf{X_u}, \mathbf{u}) \\
    &\propto p(\mathcal{D} | \Theta, \mathbf{X_u}) \times p(\Theta |
    \mathbf{X_u}) \times p(\mathcal{D} | \Theta, \mathbf{X_u}, \mathbf{u})
    \times p(\mathbf{u} | \Theta, \mathbf{X_u}) \times p(\mathcal{D} | r) \times
    p(\mathbf{r} | \Theta, \mathbf{X_u}, \mathbf{u})
  \end{split}
\]
Note that now there are only two unknown probability distributions:
$p(\mathcal{D} | \Theta, \mathbf{X_u})$ and $p(\mathcal{D} | \Theta,
\mathbf{X_u}, \mathbf{u})$, which can be computed as follows:
\begin{align*}
  p(\mathcal{D} | \Theta, \mathbf{X_u}, \mathbf{u}) = \int &p(\mathcal{D} | r) \times p(\mathbf{r} | \Theta, \mathbf{X_u}, \mathbf{u}) \, d\mathbf{r}, \\
  p(\mathcal{D} | \Theta, \mathbf{X_u}) = \iint &p(\mathcal{D} | r) \times p(\mathbf{r} | \Theta, \mathbf{X_u}, \mathbf{u}) \times p(\mathbf{u} | \Theta, \mathbf{X_u}) \, d\mathbf{u} \, d\mathbf{r}.
\end{align*}
This suggests the following form for the approximation:
\begin{equation} \label{eq:approximation}
  \approximation = q(\Theta) \times q(\mathbf{u} | \Theta) \times q(\mathbf{r}
  | \Theta, \mathbf{u}).
\end{equation}

%where $q(\Theta)$ can be of the same form as $p(\Theta | \mathbf{X_u})$,
%$q(\mathbf{u} | \Theta) = \mathcal{N}(\mathbf{0}, \bm\Sigma_{\mathbf{u}})$, and
%$q(\mathbf{r} | \Theta, \mathbf{u}) = \mathcal{N}(\bm\mu_{\mathbf{r}},
%\bm\Sigma_{\mathbf{r}})$.

@@

First, as is common in the literature for applying VI to GPs
\cite{DBLP:conf/nips/ChengB17,DBLP:conf/uai/HensmanFL13,DBLP:conf/icml/HoangHL15,DBLP:journals/jmlr/Titsias09},
we simply set
\begin{equation}
  q(\mathbf{r} | \Theta, \mathbf{u}) = p(\mathbf{r} | \Theta, \mathbf{X_u},
  \mathbf{u}).
\end{equation}
We can make a similarly justified choice for $q(\mathbf{u} | \Theta)$:
\begin{equation}
  q(\mathbf{u} | \Theta) = q(\mathbf{u}) = \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})
\end{equation}
for some variational parameters $\mathbf{m} \in \mathbb{R}^{m}$ and a positive
semi-definite matrix $\mathbf{S} \in \mathbb{R}^{m \times m}$.
\cite{DBLP:conf/nips/ChengB17,DBLP:journals/jmlr/HensmanDS17,DBLP:conf/aaai/HoangHL17}.

In order to have a reasonable way of calculating/approximating the ELBO (see
Section \ref{vi_algs}), we need to have an approximating distribution for
$\Theta$, but unfortunately all the papers in Table \ref{table:approximations}
either fix it or treat it as a variational parameter. Hence we make our first
assumption without justification from previous literature:
\begin{equation} \label{eq:theta}
  q(\Theta) = \prod_{i=0}^dq(\lambda_i).
\end{equation}
We want to restrict $\lambda_0$ to be positive so that $k_\Theta$ would produce
non-negative values and not become trivial. Similarly, we want that $\lambda_i
\ge 0$ for $i = 1, \dots, d$ so that $\bm\Lambda$ is a positive-definite matrix.
Considering the possible distributions for all $d + 1$ variables, we would like
the mean to be flexible (i.e., not tied to zero, like with the exponential
distribution), and the tails to converge to zero as the value of the random
variable moves away from the mean. We might want to support some right skew, but
the distribution should be close to symmetric with at least some parameter
values. This limits our choice of distributions quite significantly, and we
decide to go with the gamma distribution as it is fairly flexible and commonly
used \cite{hogg2018introduction}. We then define the probability density
functions as
\begin{equation} \label{eq:zeta}
  q(\lambda_i) = \Gamma(\lambda_i; \alpha_i, \beta_i) =
  \frac{\beta_i^{\alpha_i}}{\Gamma(\alpha_i)}\lambda_i^{\alpha_i - 1}e^{-\beta_i\lambda_i}, \quad i = 0, \dots, d,
\end{equation}
where $\alpha_i > 0$ and $\beta_i > 0$ are parameters of the distribution, and
$\Gamma(\cdot)$ is the gamma function. This gives us our
vector of variational parameters $\bm\nu = (\mathbf{m}, \mathbf{S}, \alpha_0,
\beta_0, \dots, \alpha_d, \beta_d)^\intercal$.

\subsection{Evidence Lower Bound}
% TODO: remove/add niu from/to q everywhere

@@

>>>>

\newcommand{\pfull}{p(\mathcal{D}, \bm\lambda, \mathbf{X_u}, \mathbf{u}, \mathbf{r})}
\newcommand{\pfullS}{p(\mathcal{D}, \bm\lambda_s, \mathbf{X_u}, \mathbf{u}_s, \mathbf{r}_s)}
\newcommand{\posterior}{p(\bm\lambda, \mathbf{u}, \mathbf{r} | \mathcal{D}, \mathbf{X_u})}
\newcommand{\approximation}{q_{\bm\nu}(\bm\lambda, \mathbf{u}, \mathbf{r})}
\newcommand{\approximationS}{q_{\bm\nu}(\bm\lambda_s, \mathbf{u}_s, \mathbf{r}_s)}

\newcommand{\Eq}{\mathbb{E}_{(\bm\lambda, \mathbf{u}, \mathbf{r}) \sim \approximation}}

@@

\section{Introduction}
%briefly explain the context of the project problem
% TODO: introduce a data set
% TODO: example application for motivation (why would the world be better)
% TODO: introduce research questions

@@

In \emph{inverse reinforcement learning}, one is presented with an MDP without a
reward function $\mathcal{M} \setminus \{ r \}$ and a set of expert
demonstrations $\mathcal{D} = \{ \zeta_i \}_{i=1}^N$, where each demonstration
$\zeta_i = \{ (s_{i,0}, a_{i,0}), \dots, (s_{i,T}, a_{i,T}) \}$ is a multiset of
state-action pairs representing the actions taken by the expert during a
particular recorded session. Each state is also characterised by a number of
features. The goal of IRL is then to find $r$ such that the optimal policy under
$r$
\[ \pi^* = \argmax_\pi \mathbb{E}\left[ \left. \sum_{t=0}^\infty \gamma^t r(s_t)
    \right| \pi \right] \]
matches the actions in $\mathcal{D}$.

Following previous work on GP IRL
\cite{DBLP:conf/nips/LevinePK11,DBLP:conf/uai/JinDAS17}, we use a maximum
entropy IRL model \cite{ziebart2008maximum}, under which we have that
\[ P(a | s) \propto \exp(Q_r(s, a)), \]
where
\begin{equation}
  Q_r(s, a) = r(s) + \gamma\sum_{s' \in \mathcal{S}}
  \mathcal{T}(s, a, s')V_r(s'),
\end{equation}
and $V_r(s)$ is a ``soft'' version of the Bellman backup operator, which can be
obtained by repeatedly applying the following equation until convergence:
\cite{DBLP:conf/nips/LevinePK11,supplementary_material}
\[ V_r(s) = \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \gamma\sum_{s' \in
      \mathcal{S}} \mathcal{T}(s, a, s')V_r(s') \right). \]
The likelihood of the data can then be written down as
\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}
\begin{equation} \label{pDr}
  p(\mathcal{D} | r) = \prod_{i=1}^N \prod_{t=1}^T p(a_{i,t} | s_{i,t}) = \exp\left( \sum_{i=1}^N \sum_{t=1}^T Q_r(s_{i,t}, a_{i,t}) - V_r(s_{i,t}) \right).
\end{equation}
However, a reward function learned by maximising this likelihood is not
transferable to new situations
\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}. One needs to model the
reward structure in a way that would allow reward predictions for previously
unseen states.

One way to model rewards without assumptions of linearity is with a
\emph{Gaussian process} (GP). A GP is a collection of random variables, any
finite combination of which has a joint Gaussian distribution
\cite{DBLP:books/lib/RasmussenW06}. We write $r \sim \mathcal{GP}(0,
k_{\bm\lambda})$ to say that $r$ is a GP with mean $0$ and covariance function
$k_{\bm\lambda}$, which uses a vector of hyperparameters $\bm\lambda$.
Covariance functions take two state feature vectors as input and quantify how
similar the two states are, in a sense that we would expect them to have similar
rewards.

As training a GP with $n$ data points has a time complexity of
$\mathcal{O}(n^3)$ \cite{DBLP:books/lib/RasmussenW06}, numerous approximation
methods have been suggested, many of which select a subset of data called
\emph{inducing points} and focus most of the training effort on them
\cite{DBLP:journals/corr/abs-1807-01065}. Let $\mathbf{X_u}$ be the matrix of
features at inducing states, $\mathbf{u}$ the rewards at those states, and
$\mathbf{r}$ a vector with $r(\mathcal{S})$ as elements. Then the full joint
probability distribution can be factorised as
\begin{equation} \label{full}
  \pfull = p(\mathbf{X_u}) \times p(\bm\lambda | \mathbf{X_u}) \times p(\mathbf{u}
  | \bm\lambda, \mathbf{X_u}) \times p(\mathbf{r} | \bm\lambda, \mathbf{X_u},
  \mathbf{u}) \times p(\mathcal{D} | r).
\end{equation}
Here $p(\mathbf{X_u})$ and $p(\bm\lambda | \mathbf{X_u})$ are customisable priors,
\begin{align}
  p(\mathbf{u} | \bm\lambda, \mathbf{X_u}) &= \mathcal{N}(\mathbf{u}; \mathbf{0}, \Kuu) \nonumber \\
                                       &= \frac{1}{(2\pi)^{m/2}|\Kuu|^{1/2}}\exp \left( -\frac{1}{2} \mathbf{u}^\intercal\Kuu^{-1}\mathbf{u} \right) \nonumber \\
                                       &= \exp\left(-\frac{1}{2}\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u} - \frac{1}{2}\log|\Kuu| - \frac{m}{2}\log 2\pi\right) \label{eq:normal}
\end{align}
is the GP prior \cite{DBLP:books/lib/RasmussenW06}, where $m \in \mathbb{N}$ is
the number of inducing points. The GP posterior is a
multivariate Gaussian \cite{DBLP:conf/nips/LevinePK11}
\begin{equation} \label{eq:r}
  p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) =
  \mathcal{N}(\mathbf{r}; \Kru^\intercal\Kuu^{-1}\mathbf{u}, \Krr - \Kru^\intercal\Kuu^{-1}\Kru),
\end{equation}
and $p(\mathcal{D} | r)$ is as in \eqref{pDr}. The matrices such as
$\Kru$ are called \emph{covariance matrices} and are defined as
$[\Kru]_{i,j} = k_{\bm\lambda}(\mathbf{x}_{\mathbf{r},i},
\mathbf{x}_{\mathbf{u},j})$, where $\mathbf{x}_{\mathbf{r},i}$ and
$\mathbf{x}_{\mathbf{u},j}$ denote feature vectors for the $i$th state in
$\mathcal{S}$ and the $j$th state in $\mathbf{X_u}$, respectively
\cite{DBLP:conf/uai/JinDAS17}.

% TODO: where does the lower bound come from?
Given this model, data $\mathcal{D}$, and inducing feature matrix
$\mathbf{X_u}$, our goal is then to find optimal values of hyperparameters
$\bm\lambda$, inducing rewards $\mathbf{u}$, and the reward function $r$. While
the previous paper that considered this IRL model computed maximum likelihood
estimates for $\bm\lambda$ and $\mathbf{u}$, and made an assumption that
$\mathbf{r}$ in \eqref{eq:r} has zero variance \cite{DBLP:conf/nips/LevinePK11},
we aim to avoid this assumption and use variational inference to approximate the
full posterior distribution $\posterior$. \emph{Variational inference} (VI) is
an approximation technique for probability densities \cite{blei2017variational}.
Let $\approximation$ be our approximating family of probability distributions
for $\posterior$ with its own hyperparameter vector $\bm\nu$. Then it is up to
VI algorithms to optimise $\bm\nu$ in order to minimise the
\emph{Kullback-Leibler} (KL) divergence between the original probability
distribution and our approximation.  KL divergence (asymmetrically) measures how
different the two distributions are and in this case can be defined as follows
\cite{blei2017variational}:
\[ \DKL(\approximation || \posterior) = \Eq \left[ \log
  \frac{\approximation}{\posterior} \right]. \]
Since KL divergence is typically hard to compute, instead of minimising it, VI
typically tries to maximise the \emph{evidence lower bound} (ELBO) defined as
\cite{DBLP:books/lib/Bishop07,blei2017variational}
\begin{equation} \label{eq:elbo}
  \begin{split}
    \mathcal{L}(\bm\nu) &= \Eq \left[ \log \frac{\pfull}{\approximation}
    \right] \\
    &= \iiint \approximation \log
    \frac{\pfull}{\approximation}\,d\bm\lambda\,d\mathbf{u}\,d\mathbf{r}.
  \end{split}
\end{equation}

@@

\begin{table}
  \centering
  \begin{tabular}{llllll}
    \toprule
    Authors, year & Inducing points & Hyperparameters & Complexity \\
    \midrule
    Titsias, 2009 \cite{DBLP:journals/jmlr/Titsias09} & variational & variational & $\mathcal{O}(nm^2)$ \\
    Hensman et al., 2013 \cite{DBLP:conf/uai/HensmanFL13} & fixed & variational & $\mathcal{O}(m^3)$ \\
    Gal et al., 2014 \cite{DBLP:conf/nips/GalWR14} & variational & variational & $\mathcal{O}(nm^2)$ \\
    Hoang et al., 2015 \cite{DBLP:conf/icml/HoangHL15} & fixed & fixed & $\mathcal{O}(m^3)$ \\
    Cheng and Boots, 2017 \cite{DBLP:conf/nips/ChengB17} & variational & variational & $\mathcal{O}(nm_\alpha + nm_\beta^2)$ \\
    Hensman et al., 2017 \cite{DBLP:journals/jmlr/HensmanDS17} & fixed & variational & $\mathcal{O}(nm)$ \\
    Peng et al., 2017 \cite{DBLP:conf/icml/PengZZQ17} & variational & variational & $\mathcal{O}(m^3)$ \\
    \bottomrule
  \end{tabular}
  \caption{Summary of relevant VI approximations to GPs. For both inducing
    points and hyperparameters, 'fixed' means 'chosen before the algorithm
    starts' and 'variational' means 'included amongst the variational
    parameters'. Hyperparameters $m_\alpha$ and $m_\beta$ refer to the number of
    bases used to represent the GP's mean and covariance, respectively.}
  \label{table:approximations}
\end{table}

@@

%state how you propose to solve the software development problem. Show that your
%proposed approach is feasible, but identify any risks.

In order to properly investigate the difference between variational inference
and maximum likelihood estimation for the model, we keep other parts of the
model the same. Namely, we set the covariance function to a version of the
automatic relevance detection kernel
\cite{DBLP:conf/nips/LevinePK11,neal2012bayesian}
\[ k_{\bm\lambda}(\mathbf{x}_i, \mathbf{x}_j) = \lambda_0\exp\left(
    -\frac{1}{2}(\mathbf{x}_i - \mathbf{x}_j)^\intercal\bm\Lambda(\mathbf{x}_i -
    \mathbf{x}_j) - \mathbbm{1}[i \ne j]\sigma^2\Tr[\bm\Lambda] \right), \]
where $\lambda_0$ is the overall ``scale'' factor for how similar or distant the
states are,
\[ \bm\Lambda = \diag(\lambda_1, \dots, \lambda_d) \]
is a diagonal matrix that determines how relevant each feature is (where $d$
denotes the number of features), $\mathbbm{1}$ is defined as
\[ \mathbbm{1}[b] = \begin{cases}
    1 & \mbox{if $b$ is true} \\
    0 & \mbox{otherwise},
  \end{cases} \]
and $\sigma^2$ is set to $10^{-2}/2$ (as the original paper noted that
the value makes little difference to the performance of the algorithm
\cite{DBLP:conf/nips/LevinePK11}). Our vector of hyperparameters for the
covariance function is then $\bm\lambda = (\lambda_0, \dots,
\lambda_d)^\intercal$. Similarly, we keep the expression for the prior of
$\bm\lambda$:
\begin{equation} \label{eq:theta_prior}
  p(\bm\lambda | \mathbf{X_u}) = \exp\left( -\frac{1}{2}\Tr[\Kuu^{-2}] -
    \sum_{i=1}^d \log(\lambda_i + 1) \right).
\end{equation}

Next, we can rewrite the posterior by using the chain rule and Bayes' theorem in
order to get a better sense of what we are trying to approximate:
\[
  \begin{split}
    \posterior &= p(\bm\lambda | \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{u}
    | \bm\lambda, \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{r} | \bm\lambda,
    \mathbf{X_u}, \mathbf{u}, \mathcal{D}) \\
    &\propto p(\bm\lambda | \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{u} |
    \bm\lambda, \mathbf{X_u}, \mathcal{D}) \times p(\mathcal{D} | r) \times
    p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) \\
    &\propto p(\bm\lambda | \mathbf{X_u}, \mathcal{D}) \times p(\mathcal{D} |
    \bm\lambda, \mathbf{X_u}, \mathbf{u}) \times p(\mathbf{u} | \bm\lambda,
    \mathbf{X_u}) \times p(\mathcal{D} | r) \times p(\mathbf{r} | \bm\lambda,
    \mathbf{X_u}, \mathbf{u}) \\
    &\propto p(\mathcal{D} | \bm\lambda, \mathbf{X_u}) \times p(\bm\lambda |
    \mathbf{X_u}) \times p(\mathcal{D} | \bm\lambda, \mathbf{X_u}, \mathbf{u})
    \times p(\mathbf{u} | \bm\lambda, \mathbf{X_u}) \times p(\mathcal{D} | r)
    \times p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u})
  \end{split}
\]
Note that now there are only two unknown probability distributions:
$p(\mathcal{D} | \bm\lambda, \mathbf{X_u})$ and $p(\mathcal{D} | \bm\lambda,
\mathbf{X_u}, \mathbf{u})$, which can be computed as follows:
\begin{align*}
  p(\mathcal{D} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) = \int &p(\mathcal{D} | r) \times p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) \, d\mathbf{r}, \\
  p(\mathcal{D} | \bm\lambda, \mathbf{X_u}) = \iint &p(\mathcal{D} | r) \times p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) \times p(\mathbf{u} | \bm\lambda, \mathbf{X_u}) \, d\mathbf{u} \, d\mathbf{r}.
\end{align*}
This suggests the following form for the approximation:
\begin{equation} \label{eq:approximation}
  \approximation = q(\bm\lambda) \times q(\mathbf{u} | \bm\lambda) \times q(\mathbf{r}
  | \bm\lambda, \mathbf{u}).
\end{equation}

@@

First, as is common in the literature for applying VI to GPs
\cite{DBLP:conf/nips/ChengB17,DBLP:conf/uai/HensmanFL13,DBLP:conf/icml/HoangHL15,DBLP:journals/jmlr/Titsias09},
we simply set
\begin{equation}
  q(\mathbf{r} | \bm\lambda, \mathbf{u}) = p(\mathbf{r} | \bm\lambda, \mathbf{X_u},
  \mathbf{u}).
\end{equation}
We can make a similarly justified choice for $q(\mathbf{u} | \bm\lambda)$:
\begin{equation}
  q(\mathbf{u} | \bm\lambda) = q(\mathbf{u}) = \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})
\end{equation}
for some variational parameters $\mathbf{m} \in \mathbb{R}^{m}$ and a positive
semi-definite matrix $\mathbf{S} \in \mathbb{R}^{m \times m}$.
\cite{DBLP:conf/nips/ChengB17,DBLP:journals/jmlr/HensmanDS17,DBLP:conf/aaai/HoangHL17}.

In order to have a reasonable way of calculating/approximating the ELBO (see
Section \ref{vi_algs}), we need to have an approximating distribution for
$\bm\lambda$, but unfortunately all the papers in Table \ref{table:approximations}
either fix it or treat it as a variational parameter. Hence we make our first
assumption without justification from previous literature:
\begin{equation} \label{eq:theta}
  q(\bm\lambda) = \prod_{i=0}^dq(\lambda_i).
\end{equation}
We want to restrict $\lambda_0$ to be positive so that $k_{\bm\lambda}$ would
produce non-negative values and not become trivial. Similarly, we want that
$\lambda_i \ge 0$ for $i = 1, \dots, d$ so that $\bm\Lambda$ is a
positive-definite matrix. Considering the possible distributions for all $d + 1$
variables, we would like the mean to be flexible (i.e., not tied to zero, like
with the exponential distribution), and the tails to converge to zero as the
value of the random variable moves away from the mean. We might want to support
some right skew, but the distribution should be close to symmetric with at least
some parameter values. This limits our choice of distributions quite
significantly, and we decide to go with the gamma distribution as it is fairly
flexible and commonly used \cite{hogg2018introduction}. We then define the
probability density functions as
\begin{equation} \label{eq:zeta}
  q(\lambda_i) = \Gamma(\lambda_i; \alpha_i, \beta_i) =
  \frac{\beta_i^{\alpha_i}}{\Gamma(\alpha_i)}\lambda_i^{\alpha_i - 1}e^{-\beta_i\lambda_i}, \quad i = 0, \dots, d,
\end{equation}
where $\alpha_i > 0$ and $\beta_i > 0$ are parameters of the distribution, and
$\Gamma(\cdot)$ is the gamma function. This gives us our
vector of variational parameters $\bm\nu = (\mathbf{m}, \mathbf{S}, \alpha_0,
\beta_0, \dots, \alpha_d, \beta_d)^\intercal$.

\subsection{Evidence Lower Bound}

@@

1. prove all applications of the dominated convergence theorem
